---
title: "HIV Analysis"
author: "Betsy Fridman"
date: "2025-11-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(readr)
```

# Load & clean data
```{r}
hiv <- read.csv("~/Desktop/MADS Fall/Applied Linear Models/HIV Project/HIV.treatments.csv")
hiv <- 
  hiv |>
  mutate(cd4_20_difference = cd4.20-cd4,
         cd4_96_difference = cd4.96-cd4.20)
hiv <-
  hiv |>
  filter(offtreat == 0)

break_points <- c(0, 8, 366, 9999)

hiv$predays_coded <-
  cut(hiv$predays, breaks = break_points, labels = c("No", "Short", "Long"), include.lowest=TRUE)
# added levels

```


# Exploratory Data Analysis
```{r}
# Response var. 1
ggplot(data = hiv, aes(x = cd4_20_difference))+
  geom_density()+
  theme_minimal()+
  labs(x = "Difference in CD4 count (20 weeks - baseline)",
       y = "Density",
       title = "Density Plot of CD4 Cell Count Difference Between 0 and 20 Weeks")
```

```{r}
# Response var. 2
ggplot(data = hiv, aes(x = cd4_96_difference))+
  geom_density()+
  theme_minimal()+
  labs(x = "Difference in CD4 count (96 weeks - 20 weeks)",
       y = "Density",
       title = "Density Plot of CD4 Cell Count Difference Between 20 and 96 Weeks")
```

```{r}
# Covariate of interest 1
ggplot(data = hiv, aes(x = cd4))+
  geom_density()+
  theme_minimal()+
  labs(x = "CD4 count at baseline",
       y = "Density",
       title = "Density Plot of CD4 Cell Count at Baseline")


ggplot(data = hiv, aes(x = cd4))+
  geom_density()+
  theme_minimal()+
  labs(x = "CD4 count at baseline",
       y = "Density",
       title = "Density Plot of CD4 Cell Count at Baseline")
```

```{r}
# Covariate of interest 2: predays
ggplot(data = hiv, aes(x = as.factor(predays_coded)))+
  geom_bar(fill = 'dodgerblue3')+
  theme_minimal()+
  labs(x = "Time intervals indicating patient history of prior HIV therapy",
       y = "Number of patients",
       title = "Bar Plot of Patients By History of Prior HIV Therapy")
```

```{r}
# Covariate of interest 3: treat2

hiv_less_advanced <-
  hiv |>
  filter(nocd4.96 == 1)

ggplot(data = hiv_less_advanced, aes(x = as.factor(treat2)))+
  geom_bar(fill = 'dodgerblue3')+
  theme_minimal()+
  labs(x = "Categories of Drug Treatments in RCT",
       y = "Number of patients",
       title = "Distribution of Patients Across Drug Treatments")+
  scale_x_discrete(labels = c("Drug A", "Drug B", "Drugs A and B", "Drugs A and C"))
```

response: cd4.20, cd4.96, cd4

Include:
  
  age - continuous
wt - continuous
hemo - factor
drugs - factor
oprior - factor
a30 - factor
predays - do something about it, keep or bin into something like hist2
race - factor
gender - factor
symptom - factor
cd4 - continuous (mabe factor for question 3)
cd8 - continuous
treat2 - continuous


Exclude:
  homo - homosexual activity doesn't affect the response to therapy, it only increases likelihood of diagnosis
karnof - idk, man-made score, sounds good but too good to be true
aprior - we would include this if everyone wasn't a 1
hist
hist2
treat
offtreat
cd4.96, don't really have to explain
nocd4.96
cd8.20, should explain maybe
cens
days

Limitation when explaining the data or eda: we don't know why people drop out and the reason for missing cd4.96 values



In eda, look to see if there's an interaction between cd4 and treat2

Question 1, would include cd4*treat2 but that would mess up my interpretation of cd4



Question 1:
cd4.20 - cd4 ~ cd4 + others
talk about if the beta of cd4 is significant and what is the sign

Question 2: 
cd4.20 - cd4 ~ predays binned + others
talk about if the beta of predays binned is significant and what is the sign

Question 3:
cd4.96 - cd4 ~ cd4 binned into less advanced (higher cd4) and more advanced (lower cd4)
talk about the coefficient of the higher cd4 bin and if its significant

Research Question 1
```{r}
model_q1 <- lm(cd4_20_difference ~ 
                 cd4 + age + wt + 
                 as.factor(hemo) + as.factor(drugs) + 
                 as.factor(oprior) + as.factor(a30) + 
                 as.factor(predays_coded) + as.factor(race) + as.factor(gender) + 
                 as.factor(symptom) + cd8 + as.factor(treat2), data=hiv)
# Assess if `cd4` is significant in estimating `cd4.20 - cd4`
summary(model_q1)
confint(model_q1, 'cd4')
```

Research Question 2
```{r}
reduced_q2 <- lm(cd4_20_difference ~ 
                 cd4 + age + wt + # excluding predays_coded 
                 as.factor(hemo) + as.factor(drugs) + 
                 as.factor(oprior) + as.factor(a30) + 
                 as.factor(race) + as.factor(gender) +
                 as.factor(symptom) + cd8 + as.factor(treat2), data=hiv)

full_q2 <- lm(cd4_20_difference ~ 
                 as.factor(predays_coded)+ cd4 + age + wt + # including predays_coded 
                 as.factor(hemo) + as.factor(drugs) + 
                 as.factor(oprior) + as.factor(a30) + 
                 as.factor(race) + as.factor(gender) + 
                 as.factor(symptom) + cd8 + as.factor(treat2), data=hiv)

anova(reduced_q2, full_q2)
```

Research Question 3
```{r}
hiv_less_advanced <-
  hiv |>
  filter(cd4 >= 500) # "A normal range is about 500-1,500"

# Filter out observations where patient didn't make it
hiv_less_advanced <-
  hiv |>
  filter(nocd4.96 != 0) # Lost about 600 observations, n = 1342 now

reduced_q3 <- lm(cd4_96_difference ~ 
                 cd4 + age + wt + # excluding treat2
                 as.factor(hemo) + as.factor(drugs) + 
                 as.factor(oprior) + as.factor(a30) + 
                 predays + as.factor(race) + as.factor(gender) + 
                 as.factor(symptom) + cd8, data=hiv_less_advanced)

full_q3 <- lm(cd4_96_difference ~ 
                 as.factor(treat2) + cd4 + age + wt + # including treat2
                 as.factor(hemo) + as.factor(drugs) + 
                 as.factor(oprior) + as.factor(a30) + 
                 predays + as.factor(race) + as.factor(gender) + 
                 as.factor(symptom) + cd8, data=hiv_less_advanced)
# Partial F Test -- won't use
anova(reduced_q3, full_q3)

# Confidence intervals for betas
summary(full_q3)
confint(full_q3, level = 0.95)
```

```{r}
ggplot(data = hiv, aes(x = cd4, y = (cd4.96-cd4.20)))+
  geom_point()
```

# Diagnostics
```{r}
diagnostics <- function(model, alpha, samplesize) {
  
  standardr <- rstandard(model)
  fittedr <- fitted(model)
  plot(fittedr, standardr,
       xlab = "Fitted values",
       ylab = "Standardized residuals",
       title = "Standardized residuals v. fitted plot")
  abline(h = c(-2, 2))# plot 1) standardized residuals vs. fitted plot
  
  library(car)
  qqPlot(standardr, distribution = "norm") # plot 2) qqPlot of standardized residuals
  
  studentr <- rstudent(model)
  plot(fittedr, studentr,
       xlab = "Fitted values",
       ylab = "Studentized residuals",
       title = "Studentized residuals v. fitted plot")
  abline(h = c(alpha/(2*samplesize),1-(alpha/(2*samplesize)))) # plot 3) studentized residuals v. fitted plot
  
  plot(model, which = 5) # plot 4) residuals v. leverage plot
  
}
```

RQ 1
```{r}
diagnostics(model_q1, 0.95, 1363)
```

Levels:
0=Drug A alone
1=Drugs A and B
2=Drugs A and C
3=Drug B alone

Given:
- "Drug A: a survival advantage is not conferred by prolonged therapy in asymptomatic subjects"

Analysis:
- Drug A is a good baseline because it does not produce a clinical benefit.
- It is neutral for all intensive purposes, so CIs of other treatments with only positive values shows significant increase in cd cell count difference from 0 to 96 weeks. AKA, there is prolonged clinical benefit from each of the other three drug therapies: A and B, A and C, and B alone.
- Does not concur with the other statements in data description of A and C not producing a clinical benefit

Conclusion:
- Given background information and analysis, A and B is most likely to produce a prolonged benefit
- There is evidence to suggest that there is prolonged clinical benefit from treatments as a whole
- Specifically A and B because
a) "Effective treatment of HIV disease will probably require combination therapy"
b) "A combination of Drug A and Drug C did not produce a clinical benefit"
- Leaves us with A and B