---
title: "Official Project 2 Script"
author: "Betsy Fridman"
date: "2025-11-13"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo=TRUE)
library(readr)
```

Read in data

```{r}
hiv <- read.csv("~/Desktop/MADS Fall/Applied Linear Models/HIV Project/HIV.treatments.csv")
hiv$hist <- as.factor(hiv$hist)
hiv$hist2 <- as.factor(hiv$hist2)
hiv$oprior <- as.factor(hiv$oprior)
hiv$aprior <- as.factor(hiv$aprior)
```

Packages

```{r}
library(tidyverse)
```

96 week data

```{r}
sum(is.na(hiv$cd4.96))
sum(hiv$nocd4.96 == 0, na.rm = TRUE)
```

The counts of missing individuals at 96 weeks is the same across the two
statistics.

Check assumption

```{r}
range(hiv$cd4)
```

Assumption is not true. The data set says values for CD4 at baseline are
between 200-500. Here, the values for CD4 range from 0-1199.

# Professor's pointers

Predays and hist

```{r}
ggplot(data = hiv,
       aes(x = hist, y = predays))+
  geom_boxplot() +
  labs(x = "HIV therapy prior to history",
       y = "Number of days of previously received HIV therapy",
       title = "Days of previously received HIV therapy, if received")+
  theme_minimal()
```

Strong relationship. Measure the same thing. May want to choose one of
two to include in the model.

Predays and hist2

```{r}
ggplot(data = hiv,
       aes(x = hist2, y = predays))+
  geom_boxplot()+
  labs(x = "HIV therapy prior to history",
       y = "Number of days of previously received HIV therapy",
       title = "Days of previously received HIV therapy, if received")
```

How is it possible for categories to overlap in predays?

```{r}
range(hiv$predays[hiv$hist2 == 3])
```

Range of predays for hist2 1st category: 0-7 Range of predays for hist2
1st category: 10-2851 Range of predays for hist2 3rd category: 20-2500

Measure slightly different things (or data error) I would include
predays and hist2 both in the model

hist and hist2

```{r}
table(hiv$hist, hiv$hist2)
```

This checks out. Only need one of hist and hist2 in the model.

hist, oprior, and aprior

```{r}
table(hiv$oprior, hiv$aprior)
table(hiv$hist, hiv$oprior)
table(hiv$hist, hiv$aprior)  # sum of prior table across all columns
```

# Other drugs vs. just drug A prior history table

Everyone used drug A prior to treatment initiation. 47 individuals also
used other drugs prior to treatment initiation.

# History of prior drug therapy vs. other drugs beyond drug A table

886 listed as "no HIV therapy history prior to study" did not use drugs
outside of Drug A prior to study \*\* Problem: Why are patients listed
as "no HIV therapy prior to study" if everyone used drug A prior to
treatment initiation? \*\* 47/2139 patients listed as having HIV therapy
prior to the study used drugs outside of just drug A

# History of prior drug therapy vs. just used drug A therapy

Most patients have prior HIV drug therapy history
